---
title       : Biostats and Epidemiology
subtitle    : Review for USMLE Step 1
author      : Austin Meyer
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

## Roadmap

 * Biostatistics - 20 minutes
  + Accuracy versus Precision - 2 minutes
  + Statistical Inference - 3 minutes
  + Distributions - 5 minutes
  + Hypothesis Testing - 10 minutes
 * Epidemiology - 40 minutes
  + Types of Prevention - 2 minutes
  + Types of Outbreaks - 2 minutes
  + Measures of Morbidity and Mortality - 5 minutes
  + Validity and Reliability of Clinical Tests - 6 minutes
  + Measures of Risk - 10
  + Epidemiological Biases - 5 minutes
  + Types of Studies - 10 minutes

--- &vcenter

<div class="centered"><font size="7">Biostatistics</font size></div>

--- &vcenter

## Accuracy is closeness and Precision is repeatability

 * Generalizability
  - How applicable is result to general population
  
 * P-value
  - Probability of finding a value this extreme by random chance
  
 * Confidence Interval
  - Interval over which real population number is found with a specified probability

--- &vcenter

## Use these terms to describe levels of belief

--- &vcenter

<div class="centered"><font size="7">Ways to describe distributions of data</font size></div>

--- &twocolumn

## Statistical distributions have invariant properties

***=left
```{r, message=FALSE, warning=FALSE, echo=FALSE}
library(ggplot2)
library(cowplot)
norm <- rnorm(10000, sd = 1)
ggplot(data.frame(data = norm), aes(data)) + 
  geom_histogram(binwidth = 0.2) + 
  xlim(-10, 10) + 
  ylim(0, 950) +
  ggtitle("Normal Distribution") + 
  labs(x="", y="") +
  annotate("text", x = -10, y = 950, label = "Mean = Median = Mode", size=6, hjust = 0) +
  annotate("text", x = -10, y = 900, label = "95% CI = Mean +- 1.96 SD", size=6, hjust = 0)
```

***=right
```{r, message=FALSE, warning=FALSE, echo=FALSE}
library(ggplot2)
library(cowplot)
studentt <- rt(10000, df=2)
ggplot(data.frame(data = studentt), aes(data)) + 
  geom_histogram(binwidth = 0.2)  + 
  xlim(-10, 10) + 
  ylim(0, 950) +
  ggtitle("Student t Distribution") + 
  labs(x="", y="") +
  annotate("text", x = -10, y = 950, label = "Mean = Median = Mode", size=6, hjust = 0) +
  annotate("text", x = -10, y = 900, label = "95% CI depends on degrees of freedom", size=6, hjust = 0)
```

--- &twocolumn

## Real distributions can have one or multiple peaks

***=left
```{r, message=FALSE, warning=FALSE, echo=FALSE}
library(ggplot2)
library(cowplot)
norm <- rnorm(10000, mean = 5, sd = 1)
ggplot(data.frame(data = norm), aes(data)) + 
  geom_histogram(binwidth = 0.1) + 
  xlim(0, 10) + 
  ylim(0, 500) +
  ggtitle("Unimodal Distribution") + 
  labs(x="", y="")
```


***=right
```{r, message=FALSE, warning=FALSE, echo=FALSE}
library(ggplot2)
library(cowplot)
norm1 <- rnorm(5000, mean = 3, sd = 0.8)
norm2 <- rnorm(5000, mean = 6, sd = 0.8)
ggplot(data.frame(data = c(norm1, norm2)), aes(data)) + 
  geom_histogram(binwidth = 0.1) + 
  xlim(0, 10) + 
  ylim(0, 500) +
  ggtitle("Bimodal Distribution") + 
  labs(x="", y="")
```

--- &twocolumn

## Skew describes the direction of the tail

***=left
```{r, message=FALSE, warning=FALSE, echo=FALSE}
library(ggplot2)
library(cowplot)
Mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}
r.skew <- rchisq(100000, df = 5)
ggplot(data.frame(data = r.skew), aes(data)) +
  geom_histogram(binwidth = 0.4, fill = 'lightgray') + 
  geom_vline(xintercept = mean(r.skew), color = 'darkred', lwd = 1.25) +
  geom_vline(xintercept = median(r.skew), color = 'darkblue', lwd = 1.25) +
  geom_vline(xintercept = Mode(round(r.skew, digits = 2)), color = 'darkgreen', lwd = 1.25) +
  xlim(0, 30) + 
  ylim(0, 6500) +
  ggtitle("Right Skewed") + 
  labs(x="", y="")
```


***=right
```{r, message=FALSE, warning=FALSE, echo=FALSE}
library(ggplot2)
library(cowplot)
Mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}
l.skew <- 30 - rchisq(100000, df = 5)
ggplot(data.frame(data = l.skew), aes(data)) + 
  geom_histogram(binwidth = 0.4, fill = 'lightgray') + 
  geom_vline(xintercept = mean(l.skew), color = 'darkred', lwd = 1.25) +
  geom_vline(xintercept = median(l.skew), color = 'darkblue', lwd = 1.25) +
  geom_vline(xintercept = Mode(round(l.skew, digits = 2)), color = 'darkgreen', lwd = 1.25) +
  xlim(0, 30) + 
  ylim(0, 6500) +
  ggtitle("Left Skewed") + 
  labs(x="", y="")
```

--- &vcenter

<div class="centered"><font size="7">Understanding hypothesis testing</font size></div>

--- &vcenter

<div class="centered"><font size="7">Epidemiology</font size></div>

---





