---
title       : Biostats and Epidemiology
subtitle    : Review for USMLE Step 1
author      : Austin Meyer
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

## Roadmap

 * Biostatistics - 20 minutes
  + Accuracy versus Precision - 2 minutes
  + Statistical Inference - 3 minutes
  + Distributions - 5 minutes
  + Hypothesis Testing - 10 minutes
 * Epidemiology - 40 minutes
  + Types of Prevention - 2 minutes
  + Types of Outbreaks - 2 minutes
  + Measures of Morbidity and Mortality - 5 minutes
  + Validity and Reliability of Clinical Tests - 6 minutes
  + Measures of Risk - 10
  + Epidemiological Biases - 5 minutes
  + Types of Studies - 10 minutes

--- &vcenter

<div class="centered"><font size="7">Biostatistics</font size></div>

---

## Accuracy is closeness and Precision is repeatability

---

## Use these terms to describe levels of belief

--- &vcenter

<div class="centered"><font size="7">Ways to describe distributions of data</font size></div>

--- &twocolumn

## Statistical distributions have invariant properties

***=left
```{r, message=FALSE, warning=FALSE, echo=FALSE}
library(ggplot2)
library(cowplot)
norm <- rnorm(10000)
ggplot(data.frame(data = norm), aes(data)) + 
  geom_histogram(binwidth = 0.1) + 
  xlim(-10, 10) + 
  ylim(0, 450) +
  ggtitle("Normal Distribution") + 
  labs(x="", y="") +
  annotate("text", x = -10, y = 450, label = "Mean = Median = Mode", size=5, hjust = 0) +
  annotate("text", x = -10, y = 430, label = "95% CI is Mean +- 1.96 SD", size=5, hjust = 0)
```


***=right
```{r, message=FALSE, warning=FALSE, echo=FALSE}
library(ggplot2)
library(cowplot)
studentt <- rt(10000, df=2)
ggplot(data.frame(data = studentt), aes(data)) + 
  geom_histogram(binwidth = 0.1)  + 
  xlim(-10, 10) + 
  ylim(0, 450) +
  ggtitle("Student t Distribution") + 
  labs(x="", y="") +
  annotate("text", x = -10, y = 450, label = "Mean = Median = Mode", size=5, hjust = 0) +
  annotate("text", x = -10, y = 430, label = "95% CI depends on degrees of freedom", size=5, hjust = 0)
```

---

## Real distributions can have one or multiple peaks

---

## Skew describes the direction of the tail

--- &vcenter

<div class="centered"><font size="7">Understanding hypothesis testing</font size></div>

--- &vcenter

<div class="centered"><font size="7">Epidemiology</font size></div>

---





